---
title: "131-hw-2"
author: "Tonia Wu"
date: "4/8/2022"
output: pdf_document
---

```{r setup, echo=FALSE}
library(tidyverse)
library(tidymodels)
library(ggplot2)
library(yardstick)
library(dplyr)
abalone <- read.csv('C:\\Users\\me\\Downloads\\homework-2\\homework-2\\data\\abalone.csv')
```

# Q1

Looks like we have a right-skewed distribution with a mean of 11.43 years.
```{r}
abalone['age'] = abalone$rings + 1.5
abalone
summary(abalone$age)
```

# Q2
The data are split with 80% in the training set and 20% in the testing set.
```{r}
# set seed
set.seed(286)

# split
aba_split <- initial_split(abalone, prop = 0.80, strata = age)
aba_train <- training(aba_split)
aba_test <- testing(aba_split)
```

# Q3
Since rings is directly used to calculate age and thus is not an independent variable, there is no use using it to predict age.

```{r}
aba_recipe <- recipe(age~type + longest_shell + diameter + height + whole_weight + shucked_weight + viscera_weight + shell_weight, data = aba_train) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_center(all_numeric_predictors()) %>%
  step_scale(all_numeric_predictors()) %>%
  step_interact(~ type:shucked_weight) %>%
  step_interact(~ longest_shell:diameter) %>%
  step_interact(~shucked_weight:skull_weight)
```

# Q4
```{r}
lm_model <- linear_reg() %>%
  set_engine('lm')
```

# Q5
```{r}
# set up a workflow
lm_wflow <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(aba_recipe)

# fit linear model to training set
lm_fit <- fit(lm_wflow, aba_train)

# results
lm_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

```

# Q6
The predicted age is  23.68:
```{r}
hypo_f_aba <- data.frame(type = 'F', longest_shell = 0.5, diameter = 0.1, height = 0.3, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1)

# results
predict(lm_fit, new_data = hypo_f_aba)
```

# Q7
As the r-squared value is only about 56%, our model only explains 56% of the variation in abalone age.

```{r}
# get training rmse
aba_train_rmse <- predict(lm_fit, new_data = aba_train %>% select(-age))
aba_train_rmse

# attach a column with observed ages
aba_train_rmse <- bind_cols(aba_train_rmse, aba_train %>% select(age))
aba_train_rmse %>% 
  head()

# get rmse
rmse(aba_train_rmse, truth = age, estimate = .pred)

# create metric set
aba_metrics <- metric_set(rmse, rsq, mae)
aba_metrics(aba_train_rmse, truth = age, estimate = .pred)
```
